---
title: 'Version 0.1.4: COVID dashboard'
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: cerulean
runtime: shiny
resource_files:
- www/Ubuntu.ttf
---


```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(skimr)
library(socviz)
library(covdata)
library(janitor)
library(DT)
library(ggthemes)
library(brotools)
library(scales)
library(leaflet)
library(plotly)
library(usmap)
library(spData)
library(ggmap)
library(hrbrthemes)
library(geofacet)
library(gganimate)
library(gifski)
library(rmapshaper)
library(sugarbag)
library(bbplot)
library(ggthemes)
library(Rttf2pt1)
library(extrafontdb)
library(extrafont)
```

Global COVID-19 Data (Maps)
=======================================================================

```{r global, message=FALSE, warning=FALSE, include=FALSE}
source("helpers.R")
# load Ubuntu font ----
extrafont::font_import(paths = "www/", prompt = FALSE)
# check fonts
# extrafont::fonts()
extrafont::loadfonts(device = "postscript")
```

Column {data-width=300 .bg-secondary}
-----------------------------------------------------------------------

### `r paste0("Total global confirmed cases as of ", WorldTSRecent$date)` 

```{r 01-world-total_confirmed}
valueBox(prettyNum(WorldTSRecent$confirmed_sum, big.mark = ","), color = "#B22222")
```

### `r paste0("New global cases as of ", WorldTSRecent$date)`

```{r 01-world-world_new}
valueBox(prettyNum(WorldTSRecent$`New Case`, big.mark = ","), color = "#EE4000")
```

### `r paste0("Global deaths as of ", WorldTSRecent$date)` 

```{r 01-world-total_deaths}
valueBox(prettyNum(WorldTSRecent$deaths_sum, big.mark = ","), color = "#EEE9E9")
```

### `r paste0("Global recovered cases as of ", WorldTSRecent$date)` 

```{r 01-world-total_recovered}
valueBox(prettyNum(WorldTSRecent$recovered_sum, big.mark = ","), color = "#00FF7F")
```

### `r paste0("Days since first confirmed case at ", case_no1)` 

```{r 01-world-days_passed}
valueBox(prettyNum(days_passed, big.mark = ","), color = "#FFFFF0")
```

Column {data-width=700 .tabset}
-----------------------------------------------------------------------

### Global Confirmed Cases

```{r geo_map_confirm_cases}
# create recent_day
recent_day <- max(SumRegionDateCodes$date)
# Set country boundaries as light gray
line <- list(color = toRGB("#d1d1d1"), width = 0.2)

# create geo for map options
geo <- list(
  bgcolor = "whitesmoke",
  showframe = FALSE,
  showcoastlines = FALSE,
  # this is the globe option
  projection = list(type = "orthographic"),
  resolution = "100",
  showcountries = TRUE,
  showocean = TRUE,
  showlakes = FALSE,
  showrivers = FALSE)

geo_map_confirm_cases <- plotly::plot_geo() %>%
  layout(
    geo = geo,
    paper_bgcolor = "whitesmoke",
    title = paste0("Global COVID-19 confirmed cases as of ", 
                   recent_day)) %>%
  add_trace(
    data = SumRegionDateCodes,
    z = ~Confirmed,
    color = ~Confirmed,
    colors = "Reds",
    text = ~country_region,
    locations = ~code,
    marker = list(line = line))

geo_map_confirm_cases
```

### Global New Cases

```{r geo_map_new_cases}
# create recent_day
recent_day <- max(SumRegionDateCodes$date)
# Set country boundaries as light gray
line <- list(color = toRGB("#d1d1d1"), width = 0.2)

# create geo for map options
geo <- list(
  bgcolor = "whitesmoke",
  showframe = FALSE,
  showcoastlines = FALSE,
  # this is the globe option
  projection = list(type = "orthographic"),
  resolution = "100",
  showcountries = TRUE,
  showocean = TRUE,
  showlakes = FALSE,
  showrivers = FALSE)

geo_map_new_cases <- plotly::plot_geo() %>%
  layout(
    geo = geo,
    paper_bgcolor = "whitesmoke",
    title = paste0("Global COVID-19 new cases as of ", 
                   recent_day)) %>%
  add_trace(
    data = SumRegionDateCodes,
    z = ~`New Cases`,
    color = ~`New Cases`,
    colors = "Oranges",
    text = ~country_region,
    locations = ~code,
    marker = list(line = line))

geo_map_new_cases
```

### Global Deaths

```{r geo_map_deaths}
# Set country boundaries as light gray
line <- list(color = toRGB("#d1d1d1"), width = 0.2)

# create geo for map options
# c("#B0E0E6", "#F0FFF0")
geo <- list(
  oceancolor = "whitesmoke",
  showframe = FALSE,
  showcoastlines = FALSE,
  # this is the mercator option
  projection = list(type = 'Mercator'),
  resolution = "100",
  showcountries = TRUE,
  showocean = TRUE,
  showlakes = FALSE,
  showrivers = FALSE)

geo_map_deaths <- plotly::plot_geo() %>%
  layout(
    geo = geo,
    paper_bgcolor = "whitesmoke",
    title = paste0("Global COVID-19 deaths as of ", 
                   recent_day)) %>%
  add_trace(
    data = SumRegionDateCodes,
    z = ~Deaths,
    color = ~Deaths,
    colors = "Greys",
    text = ~country_region,
    locations = ~code,
    marker = list(line = line))

geo_map_deaths
```

### Global Recovered Cases

```{r geo_map_recovered}
# Set country boundaries as light gray
line <- list(color = toRGB("#d1d1d1"), width = 0.2)

# create geo for map options
# c("#B0E0E6", "#F0FFF0")
geo <- list(
  oceancolor = "whitesmoke",
  showframe = FALSE,
  showcoastlines = FALSE,
  # this is the mercator option
  projection = list(type = 'Mercator'),
  resolution = "100",
  showcountries = TRUE,
  showocean = TRUE,
  showlakes = FALSE,
  showrivers = FALSE)

geo_map_recovered <- plotly::plot_geo() %>%
  layout(
    geo = geo,
    paper_bgcolor = "whitesmoke",
    title = paste0("World COVID-19 recovered cases as of ", 
                   recent_day)) %>%
  add_trace(
    data = SumRegionDateCodes,
    z = ~Recovered,
    color = ~Recovered,
    colors = "Greens",
    text = ~country_region,
    locations = ~code,
    marker = list(line = line))

geo_map_recovered
```

### Global Dataset 

```{r SumRegionDateCodes-datatable}
library(reactable)
data <- SumRegionDateCodes %>%
  dplyr::select(Country = country_region,
                `Country code` = code,
                 Date = date,
                 Confirmed,
                 `New Cases`,
                  Recovered,
                  Deaths) %>%
  dplyr::arrange(desc(Confirmed)) 
reactable::reactable(data,
  defaultSorted = "Confirmed",
  columns = list(
    Confirmed = colDef(
      name = "Confirmed",
      defaultSortOrder = "desc",
      format = colFormat(prefix = "")
    ),
    Country = colDef(
      name = "Country",
      defaultSortOrder = "desc",
      format = colFormat(separators = TRUE)
    ),
    Date = colDef(
      name = "Date",
      defaultSortOrder = "desc",
      format = colFormat(separators = TRUE)
      # format = colFormat(percent = TRUE, digits = 1)
    )
  )
)
```


Global COVID-19 Cases (Graphs)
=======================================================================

Column {data-width=300 .bg-secondary}
-----------------------------------------------------------------------

### `r paste0("Total global confirmed cases as of ", WorldTSRecent$date)` 

```{r 02-world-total_confirmed}
valueBox(prettyNum(WorldTSRecent$confirmed_sum, big.mark = ","), color = "#B22222")
```

### `r paste0("New global cases as of ", WorldTSRecent$date)`

```{r 02-world-world_new}
valueBox(prettyNum(WorldTSRecent$`New Case`, big.mark = ","), color = "#EE4000")
```

### `r paste0("Global deaths as of ", WorldTSRecent$date)` 

```{r 02-world-total_deaths}
valueBox(prettyNum(WorldTSRecent$deaths_sum, big.mark = ","), color = "#EEE9E9")
```

### `r paste0("Global recovered cases as of ", WorldTSRecent$date)` 

```{r 02-world-total_recovered}
valueBox(prettyNum(WorldTSRecent$recovered_sum, big.mark = ","), color = "#00FF7F")
```

### `r paste0("Days since first confirmed case at ", case_no1)` 

```{r 02-world-days_passed}
valueBox(prettyNum(days_passed, big.mark = ","), color = "#FFFFF0")
```

Column {data-width=700 .tabset}
-----------------------------------------------------------------------

### Global COVID-19 Cases (Animated)

```{r animate_world_cum_cases}
colors <- c("#B22222", # confirmed
            "gray65", # deaths
            "green4") # recovered

world_cum_cases <- WorldTSIncrementLong %>%

  ggplot2::ggplot(mapping = aes(x = date,
                  y = increment,
                  group = case,
                  color = case)) +
  
  ggplot2::geom_line(show.legend = FALSE) +
  
  ggplot2::scale_y_continuous(labels = scales::label_number_si(accuracy = 1)) +
  
  ggplot2::scale_color_manual(values = colors) +
  
  ggplot2::geom_segment(aes(xend = max(date) - 1,
                            
                            yend = increment),
                        
                             linetype = "dashed",
                        
                             size = 0.5,
                        
                             colour = "grey75",
                        
                             show.legend = FALSE) +
  # this adds the labels to the graph
  ggplot2::geom_text(aes(x = max(date) + 0.2, 
                         
                        label = case), 
                     
                     nudge_x = -9.0, 
                     
                     show.legend = FALSE, 
                     
                     hjust = 0) +
  # this adds the theme/font
  ggthemes::theme_few(base_size = 10, base_family = "Ubuntu") +
  # set the coordinates
  ggplot2::coord_cartesian(
    xlim = c(min(WorldTSIncrementLong$date),
             max(WorldTSIncrementLong$date) + 7),
    ylim = c(max(0, min(WorldTSIncrementLong$increment)),
             max(WorldTSIncrementLong$increment)),
    clip = "off") +
  
  ggplot2::theme(legend.position = c(0.1, 0.8),
                 # remove x axis title
                  axis.title.x = element_blank()) +
  # remove guides
  ggplot2::guides(size = FALSE) +
  # add the points 
  ggplot2::geom_point(aes(size = increment),
                      
             alpha = 0.7,
             
             show.legend = FALSE) +
  # add scale size 
  ggplot2::scale_size(range = c(2, 10)) +
  # this is the transition (x axis variable)
  gganimate::transition_reveal(date) +
  
  ggplot2::labs(title = "New Global COVID-19 Cases",
                subtitle = "Date: {frame_along}",
                        y = "New daily cases",
                        x = "Date")

animate_world_cum_cases <- gganimate::animate(world_cum_cases, 
                                              nframes = 150,
                                              fps = 10,
                                              rewind = TRUE,
                   renderer = gifski_renderer(loop = TRUE))

animate_world_cum_cases
```

### Global COVID-19 Cases (Cumulative)

```{r world_cum_point_chart}
# set colors
colors <- c("#B22222", # confirmed
            "gray65", # deaths
            "green4") # recovered
# font style
font_style <- list(
  family = "Ubuntu",
  size = 14,
  color = 'black')
# create base chart
world_cum_point_chart <- WorldTSDataAllDateLong %>%
  ggplot2::ggplot(aes(x = date,
                      y = cases,
                      color = status)) +
  geom_point(size = 1, alpha = 2/5) +
  scale_color_manual(values = colors) +
  scale_y_continuous(labels = scales::label_number_si(accuracy = 0.1)) +
    theme(
    plot.margin = margin(0, 0, 0, 0, "pt"),
    panel.background = element_rect(fill = "White"),
    legend.position = "left",
    axis.title = element_blank(),
    axis.ticks = element_blank()) +
  hrbrthemes::theme_ipsum_tw(plot_title_family = "Ubuntu") +
  labs(title = "Global COVID-19 Cumulative Cases",
       y = "Cases",
       x = "Date",
       color = " ")
# pass over to plotly
ggplotly(world_cum_point_chart) %>%
  plotly::layout(legend = list(orientation = "h"),
                 font = font_style)
```

US COVID-19 Data (Maps)
=======================================================================

Column {data-width=300 .bg-secondary}
-----------------------------------------------------------------------

### `r paste0("Total US confirmed cases as of ", SumUSRecentCountry$date_max)` 

```{r SumUSRecentCountry-confirmed}
valueBox(prettyNum(SumUSRecentCountry$confirmed_sum, big.mark = ","), color = "#B22222")
```

### `r paste0("New US cases as of ", SumUSRecentCountry$date_max)` 

```{r SumUSRecentCountry-new_case_sum}
valueBox(prettyNum(SumUSRecentCountry$new_case_sum, big.mark = ","), color = "#EE4000")
```

### `r paste0("US recovered cases as of ", SumUSRecentCountry$date_max)` 

```{r SumUSRecentCountry-recovered_sum}
valueBox(prettyNum(SumUSRecentCountry$recovered_sum, big.mark = ","), color = "#EEE9E9")
```

### `r paste0("US deaths as of ", SumUSRecentCountry$date_max)` 

```{r deaths_sum}
valueBox(prettyNum(SumUSRecentCountry$deaths_sum, big.mark = ","), color = "#00FF7F")
```

### `r paste0("Days since first confirmed case on ", us_first_case_day)` 

```{r us_days_passed}
valueBox(prettyNum(us_days_passed, big.mark = ","), color = "#FFFFF0")
```

Column {data-width=700 .tabset}
-----------------------------------------------------------------------

### United States Confirmed Cases

```{r plot_geo-confirmed_us}
us_map_layout <- list(
  scope = 'usa',
  lakecolor = "#3399FF",
  projection = list(type = 'albers usa'))

plot_geo() %>%
  add_trace(
    z = confirmed_us, 
    text = state.name, 
    span = I(0),
    locations = state.abb, 
    locationmode = 'USA-states') %>%
  layout(geo = us_map_layout,
         title = "Current US Confirmed Cases")
```

### United States New Cases

```{r plot_geo-new_case_us}
us_map_layout <- list(
  scope = 'usa',
  lakecolor = "#3399FF",
  projection = list(type = 'albers usa'))

plot_geo() %>%
  add_trace(
    z = new_case_us, 
    text = state.name, 
    span = I(0),
    locations = state.abb, 
    locationmode = 'USA-states') %>%
  layout(geo = us_map_layout,
         title = "New US Cases")
```

### United States Deaths

```{r plot_geo-deaths_us}
us_map_layout <- list(
  scope = 'usa',
  lakecolor = "#3399FF",
  projection = list(type = 'albers usa'))

plot_geo() %>%
  add_trace(
    z = deaths_us, 
    text = state.name, 
    span = I(0),
    locations = state.abb, 
    locationmode = 'USA-states') %>%
  layout(geo = us_map_layout,
         title = "Current US Deaths")
```


### US Map Data 

```{r table_state}
library(reactable)
data <- SumUSDataMap %>% 
  dplyr::select(State = state,
                Date = date,
                Confirmed = confirmed_sum,
                `New Cases` = new_case_sum,
                Deaths = deaths_sum) %>%
  dplyr::arrange(desc(Confirmed)) 
reactable::reactable(data,
  defaultSorted = "Confirmed",
  columns = list(
    State = colDef(
      name = "State",
      format = colFormat(prefix = "")
    ),
    Date = colDef(
      name = "Date",
      format = colFormat(separators = TRUE)
    ),
    Confirmed = colDef(
      name = "Confirmed",
      defaultSortOrder = "desc",
      format = colFormat(separators = TRUE)
    ),
    `New Cases` = colDef(
      name = "New Cases",
      format = colFormat(separators = TRUE)
    ),
      Deaths = colDef(
      name = "Deaths",
      format = colFormat(separators = TRUE)
    )
  )
)
```


US COVID-19 Data (Graphs)
=======================================================================

Column {data-width=300 .bg-secondary}
-----------------------------------------------------------------------

### `r paste0("Total US confirmed cases as of ", SumUSRecentCountry$date_max)` 

```{r 02-SumUSRecentCountry-confirmed}
valueBox(prettyNum(SumUSRecentCountry$confirmed_sum, big.mark = ","), color = "#B22222")
```

### `r paste0("New US cases as of ", SumUSRecentCountry$date_max)` 

```{r 02-SumUSRecentCountry-new_case_sum}
valueBox(prettyNum(SumUSRecentCountry$new_case_sum, big.mark = ","), color = "#EE4000")
```

### `r paste0("US recovered cases as of ", SumUSRecentCountry$date_max)` 

```{r 02-SumUSRecentCountry-recovered_sum}
valueBox(prettyNum(SumUSRecentCountry$recovered_sum, big.mark = ","), color = "#EEE9E9")
```

### `r paste0("US deaths as of ", SumUSRecentCountry$date_max)` 

```{r 02-deaths_sum}
valueBox(prettyNum(SumUSRecentCountry$deaths_sum, big.mark = ","), color = "#00FF7F")
```


### `r paste0("Days since first confirmed case on ", us_first_case_day)` 

```{r 02-us_days_passed}
valueBox(prettyNum(us_days_passed, big.mark = ","), color = "#FFFFF0")
```


Column {data-width=700 .tabset}
-----------------------------------------------------------------------

### US New Cases (Animated)

```{r animate_us_increment}
colors <- c("#B22222", # confirmed
            "gray65", # deaths
            "green4") # recovered

us_cum_cases <- USTSDataAllIncrementLong %>%
  # format for the K y axis
  dplyr::mutate(increment = increment/1000) %>% 
  
  ggplot2::ggplot(aes(x = date,
             y = increment,
             group = case,
             color = case)) +
    # add line
  ggplot2::geom_line(show.legend = FALSE) +
  
  ggplot2::scale_color_manual(values = colors) +
  
  # add segment, no legend
    
  ggplot2::geom_segment(aes(xend = max(date) - 1,
                            
                            yend = increment),
                        
                        linetype = "dashed",
                        
                        size = 0.5,
                        
                        color = "grey75",
                        
                        show.legend = FALSE) +
    # add text, no legend
    
  ggplot2::geom_text(aes(x = max(date) + 0.2,
                         
                         label = case),
                     
                     hjust = 0,
                     
                     nudge_x = -9.0, 
                     
                     show.legend = FALSE) +
    # set theme
      ggthemes::theme_few(base_size = 10, base_family = "Ubuntu") +
    # set cartesian coordinates to min/max dates
    ggplot2::coord_cartesian(xlim = c(min(USTSDataAllIncrementLong$date),
                                    
                                    max(USTSDataAllIncrementLong$date) + 7),
                             
                           clip = "off") +
    # position the legend
  ggplot2::theme(legend.position = c(0.1, 0.8),
                 # no x axis title
                 axis.title.x = element_blank()) +
    # no guides
  ggplot2::guides(size = FALSE) +
  
  ggplot2::geom_point(aes(size = increment),
                      
                      alpha = 0.7,
                      
                      show.legend = FALSE) +

  ggplot2::scale_size(range = c(2, 10)) +
    # set transition
  gganimate::transition_reveal(date) +
    # assign labs
  ggplot2::labs(title = "US COVID-19 Cases",
                subtitle = "at date: {frame_along}",
                y = "New Cases",
                color = "Status",
                x = "Date") + 
  
   ggplot2::scale_y_continuous(label = scales::unit_format(unit = "K")) 

animated_us_cum_cases <- gganimate::animate(us_cum_cases,
                                            nframes = 150,
                                            fps = 10,
                                            rewind = TRUE,
        renderer = gifski_renderer(loop = TRUE))

animated_us_cum_cases
```


### Positive Tests 

```{r geofacet-pos, fig.width=16, fig.height=10}
geofacet_pos <- PosMapCovus %>% 
    # adjust for scales on y axis
    dplyr::mutate(`positive tests` = `positive tests`/1000) %>% 
    # plot this with new adjusted positive tests
    ggplot2::ggplot(aes(x = days_elapsed, 
                        y = `positive tests`,
                        group = date)) +
  
    geom_col(alpha = 2/10, 
             linetype = 0) + 
  
    ggplot2::geom_line(data = TidyPosMapCovus, 
                       
    mapping = aes(x = days_elapsed, 
                  y = `Positive Test Value`/1000,
                  group = `Positive Test Metric`,
                  color = `Positive Test Metric`), 
                    show.legend = TRUE) +
      
    geofacet::facet_geo( ~ state, 
                         
                       grid = "us_state_grid2",
                       
                       scales = "free_y")  +
  
    ggplot2::labs(title = "US positive COVID tests (7-day rolling average)", 
                  
                  subtitle = paste0("Between ", 
                                    min(PosMapCovus$date), 
                                    " and ", 
                                    max(PosMapCovus$date)),
                  
                  caption = "SOURCE: https://covidtracking.com/",
                  
                  y = "Positive Tests",
                  x = "Days Elapsed") + 
  
    scale_y_continuous(label = scales::unit_format(unit = "K")) +
  
    # ggplot2::theme_bw() +
  
    ggthemes::theme_tufte(base_size = 10, base_family = "Ubuntu") +
  
    ggplot2::theme(axis.text.x = element_text(angle = 315),
                 legend.position = "top") 

  

geofacet_pos
```

### US Cases 

```{r TidyCovCaseData, fig.width=16, fig.height=10}
geofacet_cases <- TidyCovCaseData %>% 
        # adjust for y scale formatting (`K`)
        dplyr::mutate(`Cases Value` = `Cases Value`/1000) %>% 
        
        ggplot2::ggplot(aes(x = days_elapsed, 
                            
                           y = `Cases Value`, 
                  
                           group = `Cases Key`,
                  
                           color = `Cases Key`)) +
  
    ggplot2::geom_line(show.legend = TRUE) +
          
    geofacet::facet_geo( ~ state, 
                       grid = "us_state_grid2",
                       scales = "free_y")  +
  
    ggplot2::labs(title = "US COVID Cases", 
                  
                  subtitle = paste0("Between ", 
                                    min(TidyCovCaseData$date), 
                                    " and ", 
                                    max(TidyCovCaseData$date)),
                  
caption = "https://covidtracking.com | https://github.com/nytimes/covid-19-data",

                  y = "New Cases",
                  x = "Days Elapsed") + 
          
    ggplot2::scale_y_continuous(label = scales::unit_format(unit = "K")) +
  
    ggthemes::theme_tufte(base_size = 10, base_family = "Ubuntu") +
  
    ggplot2::theme(axis.text.x = element_text(angle = 315),
                 legend.position = "top") 
  
geofacet_cases
```

### US Deaths 

```{r geofacet_deaths, fig.width=16, fig.height=10}
geofacet_deaths <- TidyCovDeathData %>%
  
        ggplot2::ggplot(aes(x = days_elapsed, 
                          
                           y = `Death Value`, 
                          
                           group = `Death Key`,
                          
                           color = `Death Key`)) +
  
        ggplot2::geom_line(show.legend = TRUE) +
  
        geofacet::facet_geo( ~ state, 
                             
                             grid = "us_state_grid2", 
                             scales = "free_y")  +
  
ggplot2::labs(title = "US COVID deaths", 
              
              subtitle = paste0("Between ", 
                                
                                min(DeathsMapCovus$date), 
                                " and ", 
                                max(DeathsMapCovus$date)),
              
caption = "https://covidtracking.com | https://github.com/nytimes/covid-19-data",
              y = "New Positive Tests",
              x = "Days Elapsed") + 
      
       ggthemes::theme_tufte(base_size = 10, base_family = "Ubuntu") +
      
       ggplot2::theme(axis.text.x = element_text(angle = 315),
                     legend.position = "top") 
 
geofacet_deaths
```

Sources
=======================================================================

**Data Sources:**

1. The time series data comes from the Johns Hopkins University Center for Systems Science and Engineering (JHU CSSE) [COVID-19 dashboard data](https://github.com/CSSEGISandData/COVID-19). You can find the time series files [here](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series). These files are updated daily with new cases, recovered, and deaths.

2. The [country codes](https://github.com/datasets/gdp) come from the GDP country codes in 2016. Will be updated as needed. 

3. The [covdata](https://kjhealy.github.io/covdata/index.html) package from [Kieran Healy](https://kieranhealy.org/) contains data from the NYT database, COVID tracking project, and others. 





