---
title: "Covid - CA flexdashboard (documentation)"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---


```{r setup, echo=FALSE, cache=FALSE}
library(flexdashboard)
library(readr)
library(leaflet)
library(DT)
library(tidyverse)
library(lubridate)
library(plotly)
library(usmap)
library(spData)
library(ggmap)
library(hrbrthemes)
library(geofacet)
library(socviz)
# figs folder
fs::dir_create("figs/")
# data folder
fs::dir_create("data/")
# docs folder
fs::dir_create("docs/")
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  prompt = FALSE,
  tidy = FALSE,
  fig.width = 10,
  fig.height = 8,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.path = "figs/"
)
knitr::opts_knit$set(
  width = 78
)
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 99999
)
```

# Packages

```{r packages, include=FALSE}
# dashboard
library(flexdashboard)
library(knitr)
library(DT)
library(janitor)

# theme
library(ggthemes)
library(brotools)
library(hrbrthemes)

# data wrangling
library(tidyverse)
library(lubridate)
library(socviz)

# data visualization
library(plotly)
library(gganimate)
library(gifski)
library(skimr)
# custom skim
my_skim <- skimr::skim_with(
  numeric = skimr::sfl(p25 = NULL, p75 = NULL))

# map
library(rmapshaper)
library(sugarbag)
library(geofacet)
library(covdata)
```

***

## The `covdata` package 

We'll be using data from the [covdata package](https://kieranhealy.org/blog/archives/2020/04/10/covdata-package/) by [Kieran Healy](https://kieranhealy.org/). The goal with this package is to build an set of graphs that the user can select an input ([`selectInput()`](https://shiny.rstudio.com/reference/shiny/latest/selectInput.html)) from a list of metrics, and see that metric reflected across all 50 states. 

The datasets we'll be using are `covus` and `nytcovstate`. The script below imports and wrangles these data. 

```{r 01.1-import-wrangle-geofacet}
# fs::dir_ls("code")
source("code/01.1-import-wrangle-geofacet.R")
```


## The `covus` data 

This is a tidy dataset, with a date for each day, and each metric in the `measure` variable.  

```{r Covus}
Covus %>% dplyr::glimpse()
```

A cleaner version of the `measure` variable is stored in the `measure_label` variable.

```{r Covus-measure_label}
rmarkdown::paged_table(
dplyr::distinct(Covus, measure_label))
```

```{r Covus-measure}
rmarkdown::paged_table(
dplyr::distinct(Covus, measure))
```

Now we add the necessary date variables grouped by `state`, remove the regions not included in the `geofacet`, and put these in a dataset called `MapCovus`.

```{r MapCovus}
rmarkdown::paged_table(
MapCovus %>% head())
```

```{r PosMapCovus}
rmarkdown::paged_table(
PosMapCovus %>% head())
```


```{r TidyPosMapCovus}
rmarkdown::paged_table(
TidyPosMapCovus %>% head())
```

## The `nytcovstate` data

```{r NYTCovState}
rmarkdown::paged_table(
NYTCovState %>% head())
```

## The `TidyCovDeathData` data

```{r TidyCovDeathData}
rmarkdown::paged_table(
TidyCovDeathData %>% head())
```

# Geofacets 

We've covered how to create the [`geofacet` graphs](https://hafen.github.io/geofacet/) in [this storybench post](https://www.storybench.org/how-to-calculate-a-rolling-average-in-r/). 

We will start by building a map with the `positive` tests in it (see the output below). 

```{r}
PosMapCovus %>% 
  dplyr::filter(state == "CA") %>% 
    ggplot2::ggplot(aes(x = days_elapsed, 
                        y = `positive tests`)) +
    geom_col(alpha = 1/10) + 
    ggplot2::labs(title = "California's new COVID cases", 
                  subtitle = paste0("Total COVID-19 positive tests between ",
                               min(PosMapCovus$date), 
                               " and ",
                               max(PosMapCovus$date)),
                  y = "Positive Tests", 
                  x = "Days") + 
  hrbrthemes::theme_ipsum_rc()
```

Test California with columns and lines.

```{r TidyPosMapCA}
# get California data
TidyPosMapCA <- TidyPosMapCovus %>% dplyr::filter(state == "CA")
# plot columns
ca_pos_col_plot <- PosMapCovus %>% 
  dplyr::filter(state == "CA") %>% 
    ggplot2::ggplot(aes(x = days_elapsed, 
                        y = `positive tests`,
                        group = date)) +
    geom_col(alpha = 1/10, linetype = 0, fill = "grey50") 

ca_pos_col_plot + 
  # add the lines 
    ggplot2::geom_line(data = TidyPosMapCA, 
                       mapping = aes(x = days_elapsed, 
                                     y = `Positive Test Value`, 
                                     group = `Positive Test Metric`,
                                     color = `Positive Test Metric`), 
                       show.legend = TRUE) +
    ggplot2::labs(title = "California's new COVID cases", 
                  subtitle = paste0("Total COVID-19 positive tests between ",
                               min(PosMapCovus$date), 
                               " and ",
                               max(PosMapCovus$date)),
                  y = "Positive Tests", 
                  x = "Days") + 
  hrbrthemes::theme_ipsum_rc()
```

### US Positive Tests (COVID Tracking project)

Extend this to `geofacet`. 

```{r geofacet-pos, fig.width=16, fig.height=10}
geofacet_pos <- PosMapCovus %>% 
    ggplot2::ggplot(aes(x = days_elapsed, 
                        y = `positive tests`,
                        group = date)) +
    geom_col(alpha = 2/10, 
             linetype = 0) + 
  
    ggplot2::geom_line(data = TidyPosMapCovus, 
                       mapping = aes(x = days_elapsed, 
                                     y = `Positive Test Value`, 
                                     group = `Positive Test Metric`,
                                     color = `Positive Test Metric`), 
                       show.legend = TRUE) +
    geofacet::facet_geo( ~ state, 
                       grid = "us_state_grid1",
                       scales = "free_y")  +
  
    ggplot2::labs(title = "US positive COVID tests (7-day rolling average)", 
                  subtitle = paste0("Between ", 
                                    min(PosMapCovus$date), 
                                    " and ", 
                                    max(PosMapCovus$date)),
                  y = "New Positive Tests",
                  x = "Days Elapsed") + 
  
    ggplot2::theme_bw() +
  
    ggplot2::theme(axis.text.x = element_text(angle = 315),
                 legend.position = "top") 

  

geofacet_pos

ggsave(plot = geofacet_pos, 
       filename = "figs/geofacet_pos.png", 
       device = "png", 
       dpi = "retina",
       width = 16, 
       height = 10, 
       units = "in", 
       limitsize = FALSE)
```

### US Cases (NYT vs COVID tracking project)

Now I want to compare the NYT and COVID-19 tracking datasets (positive tests vs. NYT cases). 

```{r TidyCovCaseData, fig.width=16, fig.height=10}
geofacet_cases <- TidyCovCaseData %>% 
        ggplot2::ggplot(aes(x = days_elapsed, 
                                     y = `Cases Value`, 
                                     group = `Cases Key`,
                                     color = `Cases Key`)) +
    ggplot2::geom_line(show.legend = TRUE) +
    geofacet::facet_geo( ~ state, 
                       grid = "us_state_grid1",
                       scales = "free_y")  +
  
    ggplot2::labs(title = "US COVID Cases (NYT vs. COVID tracking)", 
                  subtitle = paste0("Between ", 
                                    min(TidyCovCaseData$date), 
                                    " and ", 
                                    max(TidyCovCaseData$date)),
                  y = "New Cases",
                  x = "Days Elapsed") + 
  
    ggplot2::theme_minimal() +
  
    ggplot2::theme(axis.text.x = element_text(angle = 315),
                 legend.position = "top") 
  


geofacet_cases

ggsave(plot = geofacet_cases, 
       filename = "figs/geofacet_cases.png", 
       device = "png", 
       dpi = "retina",
       width = 16, 
       height = 10, 
       units = "in", 
       limitsize = FALSE)
```

This can be accomplished by filtering on the measure column (which we can turn into a `selectInput`). 

```{r geofacet_deaths, fig.width=16, fig.height=10}
geofacet_deaths <- TidyCovDeathData %>% 
      ggplot2::ggplot(aes(x = days_elapsed, 
                          y = `Death Value`, 
                          group = `Death Key`,
                          color = `Death Key`)) +
    ggplot2::geom_line(show.legend = TRUE) +
    geofacet::facet_geo( ~ state, 
                       grid = "us_state_grid1",
                       scales = "free_y")  +
  
    ggplot2::labs(title = "US COVID deaths (NYT vs. COVID tracking)", 
                  subtitle = paste0("Between ", 
                                    min(DeathsMapCovus$date), 
                                    " and ", 
                                    max(DeathsMapCovus$date)),
                  y = "New Positive Tests",
                  x = "Days Elapsed") + 
  
   ggplot2::theme_minimal() +
  
   ggplot2::theme(axis.text.x = element_text(angle = 315),
                 legend.position = "top") 
  
 

geofacet_deaths

ggsave(plot = geofacet_deaths, 
       filename = "figs/geofacet_deaths.png", 
       device = "png", 
       dpi = "retina",
       width = 16, 
       height = 10, 
       units = "in", 
       limitsize = FALSE)
```


# Johns Hopkins University' CSSE Data objects

The COVID-19 data objects are loaded below from various sources. Some of these are updated daily, others are pre-existing datasets for geographic locators or other variables to aid in visualizing/mapping.

## Import time series data from Github

This imports data from the [`CSSEGISandData`](https://github.com/CSSEGISandData/COVID-19). These are in the [`csse_covid_19_time_series` folder](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series). These files are updated daily.

```{r 01.0-import-csse-time-series}
# fs::dir_ls("code/")
source("code/01.0-import-csse-time-series.R")
```


The following info comes from the README in the time series repo folder:

```md
## Time series summary (csse_covid_19_time_series)

This folder contains daily time series summary tables, including confirmed, 
deaths and recovered. All data is read in from the daily case report. 
The time series tables are subject to be updated if inaccuracies are identified
in our historical data. The daily reports will not be adjusted in these 
instances to maintain a record of raw data.

Two time series tables are for the US confirmed cases and deaths, reported at
the county level. They are named `time_series_covid19_confirmed_US.csv`, 
`time_series_covid19_deaths_US.csv`, respectively.

Three time series tables are for the global confirmed cases, recovered cases 
and deaths. Australia, Canada and China are reported at the province/state 
level. Dependencies of the Netherlands, the UK, France and Denmark are listed 
under the province/state level. The US and other countries are at the country
level. The tables are renamed `time_series_covid19_confirmed_global.csv` and 
`time_series_covid19_deaths_global.csv`, and 
`time_series_covid19_recovered_global.csv`, respectively.

### Update frequency

Once a day around 23:59 (UTC).

### Deprecated warning

The files below were archived here, and will no longer be updated. With the 
release of the new data structure, we are updating our time series tables to
reflect these changes. 

Please reference  `time_series_covid19_confirmed_global.csv` and 
`time_series_covid19_deaths_global.csv` for the latest time series data.

`time_series_19-covid-Confirmed.csv`
`time_series_19-covid-Deaths.csv`
`time_series_19-covid-Recovered.csv`
```

All of the raw data files are in this list

```{r covid_raw_data_files}
names(covid_raw_data_files)
```

I've dropped each datasets into a `raw` folder with a date stamp for safe keeping. 

```{r verify-raw-data}
fs::dir_tree(paste0("data/raw/", 
                    base::noquote(lubridate::today())), 
             recurse = FALSE)
```

## Wrangling CSSE time series data 

The code chunk below runs the script for the wrangling steps to create the data visualizations using the time series CSSE data.

```{r 02.0-wrangle-csse-time-series}
# fs::dir_ls("code/")
source("code/02.0-wrangle-csse-time-series.R")
```

These have been moved into the `processed` folder. 

```{r verify-processed-data}
fs::dir_tree(paste0("data/processed/", 
                    base::noquote(lubridate::today())), 
             recurse = FALSE)
```




```{r}
# This example modifies code from Hadley Wickham (https://gist.github.com/hadley/233134)
# It also uses data from Nathan Yau's flowingdata site (http://flowingdata.com/)
unemp <- read.csv("http://datasets.flowingdata.com/unemployment09.csv")
names(unemp) <- c("id", "state_fips", "county_fips", "name", "year", 
                  "?", "?", "?", "rate")
unemp$county <- tolower(gsub(" County, [A-Z]{2}", "", unemp$name))
unemp$state <- gsub("^.*([A-Z]{2}).*$", "\\1", unemp$name)
county_df <- map_data("county")
names(county_df) <- c("long", "lat", "group", "order", "state_name", "county")
county_df$state <- state.abb[match(county_df$state_name, tolower(state.name))]
county_df$state_name <- NULL
state_df <- map_data("state")
choropleth <- merge(county_df, unemp, by = c("state", "county"))
# sort by order
choropleth <- choropleth[order(choropleth$order), ]
# calculate rate_d
choropleth$rate_d <- cut(choropleth$rate, breaks = c(seq(0, 10, by = 2), 35))
# provide a custom tooltip to plotly with the county name and actual rate
choropleth$text <- with(choropleth, paste0("County: ", name, "Rate: ", rate))
choropleth <- choropleth %>% dplyr::select(long, lat, group, rate_d, text)
# this map takes in two datasets, choropleth and state_df
str(choropleth)
str(state_df)
```


```{r}
p <- choropleth %>% ggplot(aes(x = long, 
                               y = lat, 
                               group = group)) +
  geom_polygon(aes(fill = rate_d, 
                   text = text), 
                colour = alpha("white", 1/2), 
                              size = 0.2) + 
  geom_polygon(data = state_df, 
               colour = "white", 
               fill = NA) +
  scale_fill_brewer(palette = "PuRd") + 
  theme_void()
# just show the text aesthetic in the tooltip
ggplotly(p, tooltip = "text")
```


```{r}
density <- state.x77[, "Population"] / state.x77[, "Area"]

g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  lakecolor = toRGB('white')
)

plot_geo() %>%
  add_trace(
    z = ~density, text = state.name, span = I(0),
    locations = state.abb, locationmode = 'USA-states'
  ) %>%
  layout(geo = g)
```