---
title: "Covid - CA flexdashboard (documentation)"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---


```{r setup, echo=FALSE, cache=FALSE}
library(flexdashboard)
library(readr)
library(leaflet)
library(DT)
library(tidyverse)
library(lubridate)
library(plotly)
library(usmap)
library(spData)
library(ggmap)
library(hrbrthemes)
library(geofacet)
library(socviz)
# figs folder
fs::dir_create("figs/")
# data folder
fs::dir_create("data/")
# docs folder
fs::dir_create("docs/")
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  prompt = FALSE,
  tidy = FALSE,
  fig.width = 10,
  fig.height = 8,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.path = "figs/"
)
knitr::opts_knit$set(
  width = 78
)
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 99999
)
```

# Packages

```{r packages, include=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
# dashboard
library(flexdashboard)
library(knitr)
library(DT)
library(janitor)

# theme
library(ggthemes)
library(brotools)
library(hrbrthemes)

# data wrangling
library(tidyverse)
library(lubridate)
library(socviz)

# data visualization
library(plotly)
library(gganimate)
library(gifski)
library(skimr)
library(scales)
# custom skim
my_skim <- skimr::skim_with(
  numeric = skimr::sfl(p25 = NULL, p75 = NULL))

# map
library(rmapshaper)
library(sugarbag)
library(geofacet)
library(covdata)
```

***

## The `covdata` package 

We'll be using data from the [covdata package](https://kieranhealy.org/blog/archives/2020/04/10/covdata-package/) by [Kieran Healy](https://kieranhealy.org/). The goal with this package is to build an set of graphs that the user can select an input ([`selectInput()`](https://shiny.rstudio.com/reference/shiny/latest/selectInput.html)) from a list of metrics, and see that metric reflected across all 50 states. 

The datasets we'll be using are `covus` and `nytcovstate`. The script below imports and wrangles these data. 

```{r 01.1-import-wrangle-geofacet}
# fs::dir_ls("code")
source("code/01.1-import-wrangle-geofacet.R")
```


## The `covus` data 

This is a tidy dataset, with a date for each day, and each metric in the `measure` variable.  

```{r Covus}
Covus %>% dplyr::glimpse()
```

A cleaner version of the `measure` variable is stored in the `measure_label` variable.

```{r Covus-measure_label}
rmarkdown::paged_table(
dplyr::distinct(Covus, measure_label))
```

```{r Covus-measure}
rmarkdown::paged_table(
dplyr::distinct(Covus, measure))
```

Now we add the necessary date variables grouped by `state`, remove the regions not included in the `geofacet`, and put these in a dataset called `MapCovus`.

```{r MapCovus}
rmarkdown::paged_table(
MapCovus %>% head())
```

```{r PosMapCovus}
rmarkdown::paged_table(
PosMapCovus %>% head())
```


```{r TidyPosMapCovus}
rmarkdown::paged_table(
TidyPosMapCovus %>% head())
```

## The `nytcovstate` data

```{r NYTCovState}
rmarkdown::paged_table(
NYTCovState %>% head())
```

## The `TidyCovDeathData` data

```{r TidyCovDeathData}
rmarkdown::paged_table(
TidyCovDeathData %>% head())
```

# Geofacets 

We've covered how to create the [`geofacet` graphs](https://hafen.github.io/geofacet/) in [this storybench post](https://www.storybench.org/how-to-calculate-a-rolling-average-in-r/). 

We will start by building a map with the `positive` tests in it (see the output below). 

```{r california-new-cases-COVID-TRACKING-PROJECT}
PosMapCovus %>% 
  dplyr::filter(state == "CA") %>% 
    ggplot2::ggplot(aes(x = days_elapsed, 
                        y = `positive tests`)) +
    geom_col(alpha = 1/10) + 
    ggplot2::labs(title = "California's new COVID cases", 
                  subtitle = paste0("Total COVID-19 positive tests between ",
                               min(PosMapCovus$date), 
                               " and ",
                               max(PosMapCovus$date)),
                  y = "Positive Tests", 
                  x = "Days") + 
  hrbrthemes::theme_ipsum_rc()
```

Test California with columns and lines.

```{r TidyPosMapCA}
# get California data
TidyPosMapCA <- TidyPosMapCovus %>% dplyr::filter(state == "CA")
# plot columns
ca_pos_col_plot <- PosMapCovus %>% 
  dplyr::filter(state == "CA") %>% 
    ggplot2::ggplot(aes(x = days_elapsed, 
                        y = `positive tests`,
                        group = date)) +
    geom_col(alpha = 1/10, linetype = 0, fill = "grey50") 

ca_pos_col_plot + 
  # add the lines 
    ggplot2::geom_line(data = TidyPosMapCA, 
                       mapping = aes(x = days_elapsed, 
                                     y = `Positive Test Value`, 
                                     group = `Positive Test Metric`,
                                     color = `Positive Test Metric`), 
                       show.legend = TRUE) +
    ggplot2::labs(title = "California's new COVID cases", 
                  subtitle = paste0("Total COVID-19 positive tests between ",
                               min(PosMapCovus$date), 
                               " and ",
                               max(PosMapCovus$date)),
                  y = "Positive Tests", 
                  x = "Days") + 
  hrbrthemes::theme_ipsum_rc()
```

### Positive Tests

Extend this to `geofacet` and remove the columns. Also adjust the scales. 

```{r geofacet-pos, fig.width=16, fig.height=10}
geofacet_pos <- PosMapCovus %>% 
    # adjust for scales on y axis
    dplyr::mutate(`positive tests` = `positive tests`/1000) %>% 
    # plot this with new adjusted positive tests
    ggplot2::ggplot(aes(x = days_elapsed, 
                        y = `positive tests`,
                        group = date)) +
  
    geom_col(alpha = 2/10, 
             linetype = 0) + 
  
    ggplot2::geom_line(data = TidyPosMapCovus, 
                       
    mapping = aes(x = days_elapsed, 
                  y = `Positive Test Value`/1000,
                  group = `Positive Test Metric`,
                  color = `Positive Test Metric`), 
                    show.legend = TRUE) +
      
    geofacet::facet_geo( ~ state, 
                         
                       grid = "us_state_grid2",
                       
                       scales = "free_y")  +
  
    ggplot2::labs(title = "US positive COVID tests (7-day rolling average)", 
                  
                  subtitle = paste0("Between ", 
                                    min(PosMapCovus$date), 
                                    " and ", 
                                    max(PosMapCovus$date)),
                  
                  caption = "SOURCE: https://covidtracking.com/",
                  
                  y = "New Positive Tests",
                  x = "Days Elapsed") + 
  
    scale_y_continuous(label = scales::unit_format(unit = "K")) +
  
    # ggplot2::theme_bw() +
  
    ggthemes::theme_tufte(base_size = 10, base_family = "Ubuntu") +
  
    ggplot2::theme(axis.text.x = element_text(angle = 315),
                 legend.position = "top") 

  

geofacet_pos

ggsave(plot = geofacet_pos, 
       filename = "figs/geofacet_pos.png", 
       device = "png", 
       dpi = "retina",
       width = 16, 
       height = 10, 
       units = "in", 
       limitsize = FALSE)
```

### US Cases 

Compare the `NYT` and `COVID-tracking project` datasets (positive tests vs. NYT cases). 

```{r TidyCovCaseData, fig.width=16, fig.height=10}
geofacet_cases <- TidyCovCaseData %>% 
        # adjust for y scale formatting (`K`)
        dplyr::mutate(`Cases Value` = `Cases Value`/1000) %>% 
        
        ggplot2::ggplot(aes(x = days_elapsed, 
                            
                           y = `Cases Value`, 
                  
                           group = `Cases Key`,
                  
                           color = `Cases Key`)) +
  
    ggplot2::geom_line(show.legend = TRUE) +
          
    geofacet::facet_geo( ~ state, 
                       grid = "us_state_grid2",
                       scales = "free_y")  +
  
    ggplot2::labs(title = "US COVID Cases", 
                  
                  subtitle = paste0("Between ", 
                                    min(TidyCovCaseData$date), 
                                    " and ", 
                                    max(TidyCovCaseData$date)),
                  
caption = "https://covidtracking.com | https://github.com/nytimes/covid-19-data",

                  y = "New Cases",
                  x = "Days Elapsed") + 
          
    ggplot2::scale_y_continuous(label = scales::unit_format(unit = "K")) +
  
    ggthemes::theme_tufte(base_size = 10, base_family = "Ubuntu") +
  
    ggplot2::theme(axis.text.x = element_text(angle = 315),
                 legend.position = "top") 
  


geofacet_cases

ggsave(plot = geofacet_cases, 
       filename = "figs/geofacet_cases.png", 
       device = "png", 
       dpi = "retina",
       width = 16, 
       height = 10, 
       units = "in", 
       limitsize = FALSE)
```

### US Deaths 

NYT vs. COVID tracking project deaths.

```{r geofacet_deaths, fig.width=16, fig.height=10}
geofacet_deaths <- TidyCovDeathData %>%
        # adjust for y scale formatting (`K`)
        dplyr::mutate(`Death Value` = `Death Value`/1000) %>% 
  
        ggplot2::ggplot(aes(x = days_elapsed, 
                          
                           y = `Death Value`, 
                          
                           group = `Death Key`,
                          
                           color = `Death Key`)) +
  
        ggplot2::geom_line(show.legend = TRUE) +
  
        geofacet::facet_geo( ~ state, 
                             
                             grid = "us_state_grid2", 
                             scales = "free_y")  +
  
ggplot2::labs(title = "US COVID deaths", 
              
              subtitle = paste0("Between ", 
                                
                                min(DeathsMapCovus$date), 
                                " and ", 
                                max(DeathsMapCovus$date)),
              
caption = "https://covidtracking.com | https://github.com/nytimes/covid-19-data",
              y = "New Positive Tests",
              x = "Days Elapsed") + 
  
       ggplot2::scale_y_continuous(label = scales::unit_format(unit = "K")) +
      
       ggthemes::theme_tufte(base_size = 10, base_family = "Ubuntu") +
      
       ggplot2::theme(axis.text.x = element_text(angle = 315),
                     legend.position = "top") 
  
 
geofacet_deaths

ggsave(plot = geofacet_deaths, 
       filename = "figs/geofacet_deaths.png", 
       device = "png", 
       dpi = "retina",
       width = 16, 
       height = 10, 
       units = "in", 
       limitsize = FALSE)
```


```{r}
# This example modifies code from Hadley Wickham (https://gist.github.com/hadley/233134)
# It also uses data from Nathan Yau's flowingdata site (http://flowingdata.com/)
unemp <- read.csv("http://datasets.flowingdata.com/unemployment09.csv")
names(unemp) <- c("id", "state_fips", "county_fips", "name", "year", 
                  "?", "?", "?", "rate")
unemp$county <- tolower(gsub(" County, [A-Z]{2}", "", unemp$name))
unemp$state <- gsub("^.*([A-Z]{2}).*$", "\\1", unemp$name)
county_df <- map_data("county")
names(county_df) <- c("long", "lat", "group", "order", "state_name", "county")
county_df$state <- state.abb[match(county_df$state_name, tolower(state.name))]
county_df$state_name <- NULL
state_df <- map_data("state")
choropleth <- merge(county_df, unemp, by = c("state", "county"))
# sort by order
choropleth <- choropleth[order(choropleth$order), ]
# calculate rate_d
choropleth$rate_d <- cut(choropleth$rate, breaks = c(seq(0, 10, by = 2), 35))
# provide a custom tooltip to plotly with the county name and actual rate
choropleth$text <- with(choropleth, paste0("County: ", name, "Rate: ", rate))
choropleth <- choropleth %>% dplyr::select(long, lat, group, rate_d, text)
# this map takes in two datasets, choropleth and state_df
str(choropleth)
str(state_df)
```


```{r}
p <- choropleth %>% ggplot(aes(x = long, 
                               y = lat, 
                               group = group)) +
  geom_polygon(aes(fill = rate_d, 
                   text = text), 
                colour = alpha("white", 1/2), 
                              size = 0.2) + 
  geom_polygon(data = state_df, 
               colour = "white", 
               fill = NA) +
  scale_fill_brewer(palette = "PuRd") + 
  theme_void()
# just show the text aesthetic in the tooltip
ggplotly(p, tooltip = "text")
```


```{r}
density <- state.x77[, "Population"] / state.x77[, "Area"]

g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  lakecolor = toRGB('white')
)

plot_geo() %>%
  add_trace(
    z = ~density, text = state.name, span = I(0),
    locations = state.abb, locationmode = 'USA-states'
  ) %>%
  layout(geo = g)
```