







### `SumUSRecentCountry` = `max(date)` & group by `country` summary

`dplyr::filter` to the most recent date, group the dataset by `country` and get a summary of `date`, `confirmed`, `New Case`, and `deaths`.

```{r SumUSRecentCountry}
# SumUSRecentCountry
# these are for the valueBoxes
SumUSRecentCountry <- USTSDataAll %>%
  dplyr::filter(date == max(date)) %>%
  dplyr::group_by(country) %>%
  dplyr::summarize(
    date_max = max(date),
    confirmed_sum = sum(confirmed),
    new_case_sum = sum(`New Case`),
    deaths_sum = sum(deaths))
# create USRecentCountryMaxDate
USRecentCountryMaxDate <- SumUSRecentCountry %>%
  filter(date_max == max(date_max))
```


### `WorldTSDataRecent` = `date == max(date)`

This gives me a dataset with the most recent date.

```{r WorldTSDataRecent}
WorldTSDataRecent <- WorldTSDataAllDate %>%
  filter(date == max(date))
WorldTSDataRecent %>% rmarkdown::paged_table()
```

### Get the `USCities` data

Reduce `map::world.cities` to USA cities.

```{r WorldCities}
library(maps)
WorldCities <- maps::world.cities %>%
  dplyr::select(
    city = name,
    country_region = country.etc,
    population = pop,
    lat,
    long)
USCities <- WorldCities %>%
  dplyr::filter(str_detect(string = country_region, pattern = "USA"))
```

### Join `SumRecentRegion` to `USCities` and create `SumRecentRegionCities`.

```{r SumRecentRegionCities}
SumRecentRegionCities <- SumRecentRegion %>%
  dplyr::left_join(x = .,
                   y = USCities,
                   by = "country_region") %>%
  dplyr::arrange(desc(confirmed_sum))
glimpse(SumRecentRegionCities, 78)
```
