---
title: "Version 1.2: COVID - CA flexdashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: cerulean
runtime: shiny
---


```{r setup, include=FALSE}
# dashboard
library(flexdashboard)
library(knitr)
library(DT)
library(janitor)

# theme
library(ggthemes)
library(brotools)
library(hrbrthemes)

# data wrangling
library(tidyverse)
library(lubridate)

# data visualization
library(plotly)
library(gganimate)
library(gifski)

# map
library(rmapshaper)
library(sugarbag)
```

Global COVID-19 Data
=======================================================================

```{r global, message=FALSE, warning=FALSE, include=FALSE}
##  download csse_covid_19_time_series ----------------------------------------
# import TSConfirmedRaw ----
TSConfirmedRaw <- readr::read_csv(
  file = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")

# import TSRecoveredRaw ----
TSRecoveredRaw <- readr::read_csv(
  file = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")



# import TSDeathsRaw ----
TSDeathsRaw <- readr::read_csv(
  file = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")



# import TSConfirmedUSRaw ----------------------------------------------------
TSConfirmedUSRaw <- readr::read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")


# import TSDeathsUSRaw  --------------------------------------------------------
TSDeathsUSRaw <- readr::read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv")

# create -> ConfirmedUS --------------------------------------------------------
ConfirmedUS <- TSConfirmedUSRaw %>%
  dplyr::select(-c(UID, iso2, iso3, code3, Admin2),
                `Province_State`, `Country_Region`,
                Combined_Key, FIPS, Lat, Long_) %>%
  tidyr::pivot_longer(cols = -c(`Province_State`,
                                `Country_Region`,
                                Combined_Key,
                                FIPS,
                                Lat,
                                Long_),
                      names_to = "Date",
                      values_to = "Confirmed") %>%
  dplyr::mutate(Date = lubridate::mdy(Date)) %>%
  janitor::clean_names(case = "snake")

# create -> DeathsUS  ----
DeathsUS <- TSDeathsUSRaw %>%
  dplyr::select(-c(UID, iso2, iso3, code3, Admin2),
                `Province_State`, `Country_Region`,
                Combined_Key, FIPS, Population, Lat, Long_) %>%
  tidyr::pivot_longer(cols = -c(`Province_State`,
                                `Country_Region`,
                                Combined_Key,
                                FIPS,
                                Population,
                                Lat,
                                Long_),
                      names_to = "Date",
                      values_to = "Deaths") %>%
  dplyr::mutate(Date = lubridate::mdy(Date)) %>%
  janitor::clean_names(case = "snake")
# base::dput(lubridate::intersect(x = names(ConfirmedUS),
#                                 y = names(DeathsUS)))

# create -> USTSDataAll -----
USTSDataAll <- ConfirmedUS %>%
  dplyr::left_join(DeathsUS, by = c("fips", "province_state",
                             "country_region", "lat",
                             "long", "combined_key",
                             "date"))  %>%
  dplyr::mutate("New Case" = confirmed - lag(confirmed, 1)) %>% 
  dplyr::rename(
    state = province_state,
    country = country_region)


## create -> USMapData---------------------------------------------------------
library(socviz) # for %nin%
USMapData <- USTSDataAll %>%
  filter(state %nin% c("American Samoa", "Diamond Princess",
                       "Grand Princess", "Guam",
                       "Northern Mariana Islands", "Puerto Rico",
                       "Virgin Islands")) %>%
  filter(lat != 0.00000 & long != 0.00000)

## create -> GDPRaw ---------------------------------------------------------------------------
# import gdp from https://github.com/datasets/gdp
GDPRaw <- read_csv("https://raw.githubusercontent.com/datasets/gdp/master/data/gdp.csv")


## create -> Confirmed--------------------------------------------------------------------------------
Confirmed <- TSConfirmedRaw %>%
  tidyr::pivot_longer(cols = -c(`Province/State`,
                                `Country/Region`,
                                Lat,
                                Long),
                      names_to = "Date",
                      values_to = "Confirmed") %>%
  dplyr::mutate(Date = lubridate::mdy(Date)) %>%
  janitor::clean_names(case = "snake")


## create -> Recovered--------------------------------------------------------------------------------
Recovered <- TSRecoveredRaw %>%
  tidyr::pivot_longer(cols = -c(`Province/State`,
                                `Country/Region`,
                                Lat,
                                Long),
                      names_to = "Date",
                      values_to = "Recovered") %>%
  dplyr::mutate(Date = mdy(Date)) %>%
  janitor::clean_names(case = "snake")


## create -> Deaths-----------------------------------------------------------------------------------
Deaths <- TSDeathsRaw %>%
  tidyr::pivot_longer(cols = -c(`Province/State`,
                                `Country/Region`,
                                Lat,
                                Long),
               names_to = "Date",
               values_to = "Deaths") %>%
  dplyr::mutate(Date = mdy(Date)) %>%
  janitor::clean_names(case = "snake")


## create -> WorldTSDataAll ---------------------------------------------------------------------------
WorldTSDataAll <- Confirmed %>%
  dplyr::left_join(x = ., y = Recovered,
                   by = c("province_state",
                          "country_region",
                          "lat", "long",
                          "date")) %>%
  dplyr::mutate(recovered = replace_na(recovered,
                                       replace = 0)) %>%
  dplyr::left_join(x = ., y = Deaths,
                   by = c("province_state",
                          "country_region",
                          "lat", "long",
                          "date")) %>%
  dplyr::mutate(deaths = replace_na(deaths,
                                    replace = 0))


## create -> SumRegionDate----------------------------------------------------------------------------
# recent country_region by region
SumRegionDate <- WorldTSDataAll %>%
  # group by the region and date
  dplyr::group_by(country_region, date) %>%
  dplyr::summarize(
    confirmed_sum = sum(confirmed),
    recovered_sum = sum(recovered),
    deaths_sum = sum(deaths)) %>%
  dplyr::mutate("New Case" = confirmed_sum -
                  dplyr::lag(confirmed_sum, 1)) %>%
  dplyr::filter(date == max(date))
# rename US to USA
SumRegionDate <- SumRegionDate %>%
  dplyr::mutate(country_region = case_when(
  country_region == "US" ~ "USA",
  TRUE ~ country_region))

# remove non-physical locations
SumRegionDate <- SumRegionDate %>%
  dplyr::filter(country_region != "Diamond Princess") %>%
  dplyr::filter(country_region != "MS Zaandam")


## create -> recent_day-------------------------------------------------------------------------------
# most recent day
recent_day <- max(SumRegionDate$date)


## create -> Gdp2016----------------------------------------------------------------------------------
# filter to 2016 data
Gdp2016 <- GDPRaw %>%
  dplyr::select(region = `Country Name`,
                code = `Country Code`,
                year = Year) %>%
  dplyr::filter(year == 2016)


## create -> country_region ----------------------------------------------------
Gdp2016 <- Gdp2016 %>%
  dplyr::mutate(country_region = case_when(
   stringr::str_detect(string = region, pattern = "United States") ~ "USA",
   stringr::str_detect(string = region, pattern = "Macedonia") ~ "North Macedonia",
   stringr::str_detect(string = region, pattern = "Czech Republic") ~ "Czechia",
   stringr::str_detect(string = region, pattern = "Taiwan") ~ "Taiwan*",
   stringr::str_detect(string = region, pattern = "West Bank") ~ "West Bank and Gaza",
   stringr::str_detect(string = region, pattern = "Congo, Dem. Rep.") ~ "Congo (Kinshasa)",
   stringr::str_detect(string = region, pattern = "Congo, Rep") ~ "Congo (Brazzaville)",
   stringr::str_detect(string = region, pattern = "Bahamas, The") ~ "Bahamas",
   stringr::str_detect(string = region, pattern = "Swaziland") ~ "Eswatini",
   stringr::str_detect(string = region, pattern = "Gambia, The") ~ "Gambia",
   TRUE ~ region))


## create -> SumRecentRegionCodes---------------------------------------------------------------------
SumRegionDateCodes <- SumRegionDate %>%
  dplyr::left_join(x = ., y = Gdp2016,
                   by = "country_region") %>%
  dplyr::arrange(desc(confirmed_sum))
# easier printing
SumRegionDateCodes <- SumRegionDateCodes %>% 
  dplyr::rename(`Confirmed` = confirmed_sum)

## create -> WorldTSDataAllDate ----------------------------------------------
WorldTSDataAllDate <- WorldTSDataAll %>%
  # change this to region
  dplyr::rename(region = country_region) %>%
  # group by dates
  dplyr::group_by(date) %>%
  # get summary variables
  dplyr::summarize(
    confirmed_sum = sum(confirmed),
    deaths_sum = sum(deaths),
    recovered_sum = sum(recovered)
  ) %>%
  # create new case with lag
  dplyr::mutate("New Case" = confirmed_sum - dplyr::lag(confirmed_sum, 1))


## create -> WorldTSDataAllDateLong --------------------------------------------
WorldTSDataAllDateLong <- WorldTSDataAllDate %>%
  dplyr::select(-c(`New Case`),
         `Confirmed` = confirmed_sum,
         `Deaths` = deaths_sum,
         `Recovered` = recovered_sum) %>%
  tidyr::pivot_longer(cols = -date,
               names_to = "status",
               values_to = "cases")


## create -> WorldTSIncrementLong---------------------------------------------------------------------
WorldTSIncrementLong <- WorldTSDataAllDate %>%
  dplyr::select(-`New Case`) %>%
  dplyr::mutate(
    `Confirmed` = confirmed_sum - lag(confirmed_sum, 1),
    `Deaths` = deaths_sum - lag(deaths_sum, 1),
    `Recovered` = recovered_sum - lag(recovered_sum, 1)) %>%
  dplyr::filter(date != min(date)) %>%
  tidyr::pivot_longer(cols = c(`Confirmed`, `Deaths`, `Recovered`),
               names_to = "case",
               values_to = "increment")
# WorldTSIncrementLong

## create -> WorldTSDataUS -----------------------------------------------------
WorldTSDataUS <- WorldTSDataAll %>%
  dplyr::filter(country_region == "US") %>%
  dplyr::rename(
    state = province_state,
    country = country_region) %>%
  dplyr::mutate("New Case" = confirmed - lag(confirmed, 1))

## create -> USTSDataAllIncrementLong ------------------------------------------
USTSDataAllIncrementLong <- WorldTSDataUS %>%
  dplyr::select(date, confirmed, recovered, deaths, state) %>%
  dplyr::group_by(state) %>%
  dplyr::mutate(
    confirmed_lag = confirmed - lag(confirmed, 1),
    recovered_lag = recovered - lag(recovered, 1),
    deaths_lag = deaths - lag(deaths, 1)) %>%
  dplyr::filter(date != min(date)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(date) %>%
  dplyr::summarize(
    Confirmed = sum(confirmed_lag),
    Recovered = sum(recovered_lag),
    Deaths = sum(deaths_lag)) %>%
  tidyr::pivot_longer(-date,
               names_to = "case",
               values_to = "increment")

# create valueBox data ----
WorldTSRecent <- WorldTSDataAllDate %>%
  filter(date == max(date))
# SumUSRecentCountry
# these are for the valueBoxes
SumUSRecentCountry <- WorldTSDataUS %>%
  dplyr::filter(date == max(date)) %>%
  dplyr::group_by(country) %>%
  dplyr::summarize(
    date_max = max(date),
    confirmed_sum = sum(confirmed),
    recovered_sum = sum(recovered),
    new_case_sum = sum(`New Case`),
    deaths_sum = sum(deaths))
# value boxes first tab
new_case <- WorldTSDataAll %>% dplyr::filter(confirmed > 0)
case_no1 <- min(new_case$date)
tday <- Sys.Date()
days_passed <- tday - case_no1
us_new_conf <- WorldTSDataUS %>% dplyr::filter(confirmed > 0)
us_first_case_day <- min(us_new_conf$date)
tday <- Sys.Date()
us_days_passed <- tday - us_first_case_day
```

Column {data-width=400 .bg-secondary}
-----------------------------------------------------------------------

### `r paste0("New cases as of ", WorldTSRecent$date)`

```{r world_new}
valueBox(prettyNum(WorldTSRecent$`New Case`, big.mark = ","), color = "tomato")
```

### `r paste0("Confirmed cases as of ", WorldTSRecent$date)` 

```{r total_confirmed}
valueBox(prettyNum(WorldTSRecent$confirmed_sum, big.mark = ","), color = "sandybrown")
```

### `r paste0("Recovered cases as of ", WorldTSRecent$date)` 

```{r total_recovered}
valueBox(prettyNum(WorldTSRecent$recovered_sum, big.mark = ","), color = "palegreen")
```

### `r paste0("Deaths as of ", WorldTSRecent$date)` 

```{r total_deaths}
valueBox(prettyNum(WorldTSRecent$deaths_sum, big.mark = ","), color = "gray50")
```


### `r paste0("Days since first confirmed case at ", case_no1)` 

```{r days_passed}
valueBox(prettyNum(days_passed, big.mark = ","), color = "lightgoldenrodyellow")
```

Column {data-width=600 .tabset}
-----------------------------------------------------------------------

### Global COVID-19 Confirmed Cases

```{r geo_map_confirm_cases}
recent_day <- max(SumRegionDateCodes$date)
# Set country boundaries as light gray
line <- list(color = toRGB("#d1d1d1"), width = 0.2)

# create geo for map options
geo <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = "orthographic"),
  resolution = "100",
  showcountries = TRUE,
  countrycolor = "#EEC591",
  showocean = TRUE,
  oceancolor = "#00B2EE",
  showlakes = TRUE,
  lakecolor = "#BFEFFF",
  showrivers = TRUE,
  rivercolor = "#0000FF",
  bgcolor = "#FFFFFF"
)

geo_map_confirm_cases <- plotly::plot_geo() %>%
  layout(
    geo = geo,
    paper_bgcolor = "#e8f7fc",
    title = paste0("World COVID-19 Confirmed Cases by Region at ", recent_day)
  ) %>%
  add_trace(
    data = SumRegionDateCodes,
    z = ~`Confirmed`,
    color = ~`Confirmed`,
    colors = "Oranges",
    text = ~`country_region`,
    locations = ~code,
    marker = list(line = line)
  )

geo_map_confirm_cases
```


Daily COVID-19 Cases (Global)
=======================================================================

Column {data-width=500}
-----------------------------------------------------------------------

### New Daily World COVID-19 Cumulative Cases

```{r world_cum_point_chart}
# set colors
colors <- c("red1", "gray1", "springgreen2")
# create base chart
world_cum_point_chart <- WorldTSDataAllDateLong %>%
  ggplot2::ggplot(aes(x = date,
                      y = cases,
                      color = status)) +
  ggplot2::geom_point(size = 1, alpha = 2/5) +
  ggplot2::scale_color_manual(values = colors) +
  ggplot2::scale_y_continuous(labels = scales::label_number_si(accuracy = 0.1)) +
    ggplot2::theme(
    plot.margin = margin(0, 0, 0, 0, "pt"),
    panel.background = element_rect(fill = "White"),
    legend.position = "none",
    axis.title = element_blank(),
    axis.ticks = element_blank()) +
  hrbrthemes::theme_ipsum_tw(plot_title_family = "Ubuntu") +
  ggplot2::labs(title = "World COVID-19 Cumulative Cases by Day",
       y = "Cases",
       x = "Date",
       color = " ")
# pass over to plotly
ggplotly(world_cum_point_chart) %>%
  layout(legend = list(orientation = "h"))
```

Column {data-width=500}
-----------------------------------------------------------------------

### New World COVID-19 Case Each Day

```{r animate_world_cum_cases}
colors <- c("red1", "gray1", "springgreen2")
world_cum_cases <-
  WorldTSIncrementLong %>%

  ggplot2::ggplot(mapping = aes(x = date,
                  y = increment,
                  group = case,
                  color = case)) +
  ggplot2::geom_line(show.legend = FALSE,) +
  ggplot2::scale_y_continuous(labels = scales::label_number_si(accuracy = 1)) +
  ggplot2::scale_color_manual(values = colors) +
  ggplot2::geom_segment(aes(xend = max(date) + 1,
                            yend = increment),
                             linetype = 2,
                             colour = "grey1",
                             show.legend = FALSE) +
  ggplot2::geom_text(aes(x = max(date) + 1,
                label = case),
                 show.legend = FALSE,
                 hjust = 0) +
  hrbrthemes::theme_ipsum_tw(plot_title_family = "Ubuntu") +
  # set the coordinates
  ggplot2::coord_cartesian(
    xlim = c(min(WorldTSIncrementLong$date),
             max(WorldTSIncrementLong$date) + 7),
    ylim = c(max(0, min(WorldTSIncrementLong$increment)),
             max(WorldTSIncrementLong$increment)),
    clip = "off") +
  ggplot2::theme(legend.position = c(0.1, 0.8),
        axis.title.x = element_blank()) +
  ggplot2::guides(size = FALSE) +
  ggplot2::geom_point(aes(size = increment),
             alpha = 0.7,
             show.legend = FALSE) +

  ggplot2::scale_size(range = c(2, 10)) +

  gganimate::transition_reveal(date) +
  ggplot2::labs(title = "New Global COVID-19 Cases",
      subtitle = "Date: {frame_along}",
      y = "New daily cases",
      x = "Date")
animate_world_cum_cases <- gganimate::animate(world_cum_cases, nframes = 150,
                                              fps = 10,
                   renderer = gifski_renderer(loop = TRUE))
animate_world_cum_cases
```


Daily COVID-19 Cases (United States)
=======================================================================

Column {data-width=400 .bg-secondary}
-----------------------------------------------------------------------

### `r paste0("New cases as of ", SumUSRecentCountry$date_max)` 

```{r SumUSRecentCountry-new_case_sum}
valueBox(prettyNum(SumUSRecentCountry$new_case_sum, big.mark = ","), color = "tomato")
```

### `r paste0("Confirmed cases as of ", SumUSRecentCountry$date_max)` 

```{r SumUSRecentCountry-confirmed}
valueBox(prettyNum(SumUSRecentCountry$confirmed_sum, big.mark = ","), color = "sandybrown")
```

### `r paste0("Recovered cases as of ", SumUSRecentCountry$date_max)` 

```{r SumUSRecentCountry-recovered_sum}
valueBox(prettyNum(SumUSRecentCountry$recovered_sum, big.mark = ","), color = "palegreen")
```

### `r paste0("Deaths as of ", SumUSRecentCountry$date_max)` 

```{r aus_deaths}
valueBox(prettyNum(SumUSRecentCountry$deaths_sum, big.mark = ","), color = "gray50")
```


### `r paste0("Days since first confirmed case on ", us_first_case_day)` 

```{r us_days_passed}
valueBox(prettyNum(us_days_passed, big.mark = ","), color = "#FFC300")
```

Column {data-width=600 .tabset}
-----------------------------------------------------------------------

### New US COVID-19 Cases Each Day

```{r animate_us_increment}
us_cum_cases <- USTSDataAllIncrementLong %>%
  ggplot2::ggplot(aes(x = date,
             y = increment,
             group = case,
             color = case)) +
    # add line
  ggplot2::geom_line(show.legend = FALSE) +
  ggplot2::scale_color_manual(values = c("red1", "black", "dodgerblue")) +
    # add segment, no legend
  ggplot2::geom_segment(aes(xend = max(date) + 1,
                            yend = increment),
                        linetype = 2,
                        color = "grey",
                        show.legend = FALSE) +
    # add text, no legend
  ggplot2::geom_text(aes(x = max(date) + 1,
                         label = case),
                     hjust = 0,
                     show.legend = FALSE) +
    # set theme
  hrbrthemes::theme_ipsum_tw(plot_title_family = "Ubuntu") +
    # set cartesian coordinates to min/max dates
  ggplot2::coord_cartesian(xlim = c(min(USTSDataAllIncrementLong$date),
                                    max(USTSDataAllIncrementLong$date) + 7),
                           clip = "off") +
    # position the legend
  ggplot2::theme(legend.position = "top",
                 # no x axis title
                 axis.title.x = element_blank()) +
    # no guides
  ggplot2::guides(size = FALSE) +
  ggplot2::geom_point(aes(size = increment),
                      alpha = 0.7,
                      show.legend = FALSE) +

  ggplot2::scale_size(range = c(2, 10)) +
    # set transition
  gganimate::transition_reveal(date) +
    # assign labs
  ggplot2::labs(title = "US COVID-19 Cases",
                subtitle = "at date: {frame_along}",
                y = "New Cases",
                color = "Status",
                x = "Date")

animated_us_cum_cases <- gganimate::animate(us_cum_cases,
                                            nframes = 150,
                                            fps = 10,
        renderer = gifski_renderer(loop = TRUE))
animated_us_cum_cases
```


Geofacet COVID-19 Data
=======================================================================

Column {data-width=500 .tabset}
-----------------------------------------------------------------------

```{r }

```

```{r }

```

### Europe COVID-19 Confirmed Cases

```{r }

```

### Europe COVID-19 Recovered Cases

```{r }

```

### Europe COVID-19 Deaths

```{r }

```


Column {data-width=500 .tabset}
-----------------------------------------------------------------------

### US COVID-19 Confirmed Cases

```{r ggfacet_confirmed}

```


### US COVID-19 Recovered

```{r ggfacet_recovered}

```


### US COVID-19 Deaths

```{r ggfacet_deaths}

```


Sources
=======================================================================

1. [COVID-19 data](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series)

3. [The country codes](https://github.com/datasets/gdp) 


Column {data-width=500}
-----------------------------------------------------------------------

### The Global Data 

```{r SumRegionDateCodes-datatable}
SumRegionDateCodes %>%
  dplyr::select(`Country region` = country_region,
         `Country code` = code,
         `Date` = date,
         `Confirmed`,
         `New Case`,
         `Recovered` = recovered_sum,
         `Deaths` = deaths_sum) %>%
  dplyr::arrange(desc(`Confirmed`)) %>%
  DT::datatable(
    rownames = FALSE,
    fillContainer = TRUE,
    options = list(
      bPaginate = FALSE
    )
  )
```


Column {data-width=500}
-----------------------------------------------------------------------

### The US Data 

```{r table_state}
WorldTSDataUS %>%
  select(Country = country,
         State = state,
         Date = date,
         Confirmed = confirmed,
         `New Case`,
         Recovered = recovered,
         Deaths = deaths) %>%
  arrange(desc(Confirmed)) %>%
  DT::datatable(
    rownames = FALSE,
    fillContainer = TRUE,
    options = list(
      bPaginate = FALSE
    )
  )
```

